Biotic invasions fuck with the manageability of pollination networks

### 2017-02-20 paper/supp_info_2.Rmd
Add control profiles analysis

# S2: Properties of empirical networks

@@

```{r fig-visitation-vs-deposition-vs-importance, echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 4, results = "hide", fig.cap= "**Distribution of interaction weights for the pollen deposition, visitation and pollinator importance networks**. Note that the *x* axis in the density plots have been log-transformed."}
n_driver <- meta %>%
  dplyr::inner_join(n_matched) %>% 
  dplyr::mutate(n_driver = 1 - n_matched / (n_pla + n_pol))

>>>>

# Control profiles

@Ruths2014 proposed a heuristic-based method to provide insight of what might be required to control a network.
The method is based on the idea that the reasons behind a node being classified as a driver node can be precisely identified, and that the relative contribution of these reasons can be used to characterise the control profile of the network. 
Specifically, nodes can be deemed to be driver nodes for three reasons:

1. **Because it is a source node**. Source nodes are nodes that exclusively have outgoing links. Therefore, other nodes cannot control it, but instead, it must be externally controlled. 
2. **Because of an external dilation**. Dilations occur whenever a control signal needs to branch out in order to reach all nodes in a network. 
One kind of dilation, which arises due to a surplus of sink nodes (those that only have incoming links) is called external dilation. 
This concept might be easier to understand by looking at Figure \@ref(fig:fig-SA-exp)c. 
The source node *a* branches out into nine paths, but it can only control one of them. 
We say then, that the sink nodes *d* to *k* become driver nodes because of an external dilation.
3. **Because of an internal dilation**. The remaining dilations (those that do not arise from a surplus of sink nodes) are called internal dilations. 
Node *j* in Figure \@ref(fig:fig-SA-exp)c and the driver nodes in Figure \@ref(fig:fig-matchings-bidirectional) are examples of this category.

Mathematically, the number of driver nodes $N_D$ is the sum of the number of source nodes, the number of external dilation points, and the number of internal dilation points, $N_D = N_s + N_e + N_i$. 
A network can then be characterised by the relative proportion of the three kind of controls $$\eta_s + \eta_e + \eta_i = 1,$$ where $\eta_s = N_s/N_D$, $\eta_e = N_e/N_D$, and $\eta_i = N_i/N_D$. 
These terms can be calculated with relative ease.
$N_D$ can be calculated using a maximum matching algorithm. 
$N_s$ can be calculated by inspecting the network's degree distribution.
$N_e = max(0,N_t - N_s)$, where $N_t$ is the number of sink nodes in the network.
And we can solve for $N_i$ once we have all other information.

Applying this framework to our networks directed by the asymmetry we identify that they are all external dilation dominated (Figure \@ref(fig:fig-control-profiles)).
@Ruths2014 explains that "external-dilation-dominated networks exhibited strong aspects of top-down control". 
Because of the surplus of sink nodes, an external intervention that is applied to a source node will induce a correlated response among their subordinate species. 
For example, an increase in the abundance of *Heracleum sphondylum* (Figure \@ref(fig:fig-SA-exp)) is likely to induce an increase in the abundance of the pollinators it interacts with.
Therefore it can be expected that in order to fully control the network it is not sufficient to apply interventions to the source nodes.

The control profile of our pollination networks provides additional justification to our decision to determine the importance of a species, not only by virtue of $f_D$, but also by $f_S$.
Many species that act like control sinks in our network, will be classified as a driver node, because they are necessary to *fully* control the dynamics of the community, as opposed to playing a central role influencing the abundance of other species in the community.

Using this approach also highlights the importance of appropiately selecting the direction of control (See Section 6). 
A network with bidirectional links weighted by the mutual dependences would have zero source nodes and zero sink nodes. 
Therefore, the control profile of such a network would have been completely uninformative, as it would just highlight that the proportion of internal dilations is $\eta_i = 1$.

```{r fig-control-profiles, echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 2.6, results = "hide", fig.cap= "**Control profiles of the studied networks**. Following @Ruths2014, profiles are shown in a ternary plot in which the cornders correspond to points (1, 0, 0), (0, 1, 0), and (0, 0, 1) in the three-dimensional $(\\eta_s, \\eta_e, \\eta_i)$ space such that $\\eta_s + \\eta_e + \\eta_i = 1$, and $\\eta_s, \\eta_e$, and $\\eta_i > 0$."}
n_driver <- meta %>%
  dplyr::inner_join(n_matched) %>% 
  dplyr::mutate(n_driver = 1 - n_matched / (n_pla + n_pol))

profiles <- net %>%
	plyr::ldply(function(x){
		d <- x %>% 
			keep_largest_component() %>%
			bipartite_digraph(type = "weight",
												keep = "all",
												weight.type = "asymmetry", 
												scale = F)
		E(d)$weight <- 1
		out_degree <- igraph::degree(d, mode = "out") 
    in_degree <- igraph::degree(d, mode = "in") 
    n_sources <- sum(out_degree > 0 & in_degree == 0) 
    n_sinks <- sum(out_degree == 0 & in_degree > 0) 
    n_ext_dil <- max(0,n_sinks - n_sources) 
    n_sp <- length(igraph::V(d))
    data.frame(n_sources = n_sources, n_sinks = n_sinks,  
               n_ext_dil = n_ext_dil, n_sp = n_sp)
	}) %>%
	dplyr::rename(net_name = .id) %>%
	dplyr::mutate()

profiles %<>%
	dplyr::inner_join(n_driver) %>%
	dplyr::mutate(n_driver = n_sp - n_matched,
								n_int_dil = n_driver - n_sources - n_ext_dil,
								eta_s = n_sources / n_driver,
								eta_e = n_ext_dil / n_driver,
								eta_i = n_int_dil / n_driver)

library(ggtern)

# p1 <- 
profiles %>%
 	dplyr::filter(!grepl("bal_bin", net_name)) %>%
 	ggplot(aes(x = eta_s, z = eta_e, y = eta_i)) + 
	geom_mask() +
  geom_point(fill = "gray70", shape = 21, size = 2, alpha = 0.7) + 
	coord_tern(expand = T) +
 	labs(x = "$\\eta_s$",
 			 y = "$\\eta_i$",
 			 z = "$\\eta_e$",
 			 Tarrow = "internal dilation",
 			 Rarrow = "external dilation",
 			 Larrow = "source") +
	fer_theme +
 	# tern_limit(0.2,0.2, 1) +
 	theme(panel.grid = element_line(colour = "grey90", size = 0.3),
 				plot.title = element_text(margin = margin(-14, 0, 0, 0)),
 				plot.margin = margin(t = -10),
 				plot.background = element_blank()) +
 	percent_custom("%") +
 	theme_nogrid_minor() +
 	theme_showarrows() +
 	theme_arrowlarge() +
 	theme_latex(TRUE) 
# 
# p2 <- profiles %>% 
# 	dplyr::filter(!grepl("bal_bin", net_name)) %>%
# 	ggplot(aes(eta_e, n_driver/n_sp)) +
# 	geom_point(shape = 21) +
# 	# scale_x_log10() + scale_y_log10() +
# 	geom_smooth(method = "lm", colour = "black", size = 0.5) +
# 	# scale_x_log10() +
# 	fer_theme +
# 	theme(plot.margin = margin(r = 2, t = 18, b = 21),
#  				plot.background = element_blank()) +
# 	xlab(expression(eta[s])) +
# 	ylab(expression(n[D]))
# 
# grid.arrange(grobs = list(
# 	grid::textGrob("(a)", x = 0, y = -2, gp = grid::gpar(font = 2, cex = 0.7), just = "left"),
# 	grid::textGrob("(b)", x = 0, y = -2, gp = grid::gpar(font = 2, cex = 0.7), just = "left"), 
# 	p1, p2), nrow = 2, heights = c(1,30), widths = c(1, 1.4))
```

# Properties of empirical networks

@@

```{r fig-visitation-vs-deposition-vs-importance, echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 4, results = "hide", fig.cap= "**Distribution of interaction weights for the pollen deposition, visitation and pollinator importance networks**. Note that the *x* axis in the density plots have been log-transformed."}
